A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-09-18 paper.Rmd
intro done.

It is also common to make scatterplots of residuals against the fitted values, and each of the explanatory variables, can be used to scrutinize their relationships. If there are any visually discoverable associations, the model is potentially inadequate or incorrectly specified. \textcolor{red}{It is also possible to discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.} To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity. A non-linear pattern would suggest that the model needs to have some additional non-linear terms. Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption. Statistical tests were developed to provide objective assessment, for example, of non-linear patterns [e.g. @ramsey1969tests], and heteroskedasticity  [e.g. @breusch1979simple].

@@

\textcolor{red}{However, it needs to be noted that relying purely on subjective assessment of plots can be a source of anxiety. Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a really difficult task. People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise. \cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}

>>>>

\textcolor{red}{Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships}. If there are any visually discoverable associations, the model is potentially inadequate or incorrectly specified. \textcolor{red}{We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.} To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity. A non-linear pattern would suggest that the model needs to have some additional non-linear terms. Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption. Statistical tests were developed to provide objective assessment, for example, of non-linear patterns [e.g. @ramsey1969tests], and heteroskedasticity  [e.g. @breusch1979simple].

@@

\textcolor{red}{However, relying purely on subjective assessment of a single plot can be deceiving. Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task. People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise. \cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}

### 2023-09-18 paper.tex
intro done.

It is also common to make scatterplots of residuals against the fitted
values, and each of the explanatory variables, can be used to scrutinize
their relationships. If there are any visually discoverable
associations, the model is potentially inadequate or incorrectly
specified.
\textcolor{red}{It is also possible to discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.}
To read residual plots, one looks for noticeable departures from the
model such as non-linear pattern or heteroskedasticity. A non-linear
pattern would suggest that the model needs to have some additional
non-linear terms. Heteroskedasticity suggests that the error is
dependent on the predictors, and hence violates the independence
assumption. Statistical tests were developed to provide objective
assessment, for example, of non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.

>>>>

\textcolor{red}{Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships}.
If there are any visually discoverable associations, the model is
potentially inadequate or incorrectly specified.
\textcolor{red}{We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.}
To read residual plots, one looks for noticeable departures from the
model such as non-linear pattern or heteroskedasticity. A non-linear
pattern would suggest that the model needs to have some additional
non-linear terms. Heteroskedasticity suggests that the error is
dependent on the predictors, and hence violates the independence
assumption. Statistical tests were developed to provide objective
assessment, for example, of non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.
